''' By Shubham Agarwal
    Link: https://github.com/BeAgarwal/R '''
    
Advanced Programming in R

Functions in R
-> A function accepts input arguments and produces output by executing valid R commands present n the function.
-> Function name and function names need not to be same.
-> Functions are created using the command function().

```{r}
area = function(r=10){
  a = 2 * 3.14 * r
  return (a)
}
ans = area(20)
ans
```

Function with multiple inputs and outputs.
```{r}
rectangle = function(l, b){
  ar = l * b
  pr = 2 * (l + b)
  result = list("Area"=ar, "perimeter"=pr)
  return (result)
}
ans = rectangle(10,20)
ans
```

Inline Function
```{r}
func = function(x) x^2+5*x+20
func(9)
```

Looping over Objects:
There are a few looping functions that are pretty useful when working interactively on a command line.
apply -> Apply a function over the margins of a array or matrix
lapply -> Apply a function over a list or a vector
tapply -> Apply a function over a ragged array
mapply -> Multivariate version of lapply
xxply -> plyr package

apply function:
Syntax: apply(array, margins, function,...)
```{r}
A = matrix(1:9,3,3)
A
apply(A,1,sum)  #Row
apply(A,2,sum)  #Column
```

lapply function:
Syntax: mapply(list, function,...)
```{r}
A = matrix(1:9,3,3)
B = matrix(10:18,3,3)
list1 = list(A,B)
determinant = lapply(list1,det)
determinant
```

mapply function:
Syntax: mapply(function, list1, list2,...)
```{r}
rectangle = function(l, b){
  ar = l * b
  pr = 2 * (l + b)
  result = list("Area"=ar, "perimeter"=pr)
  return (result)
}
A = list(1,2,3,4)
B = list(10,20,40,30)
determinant = mapply(rectangle, A, B)
determinant
```

tapply function:
Syntax: tapply(vector, factors, function,...)
```{r}
Id = c(1,1,1,1,2,2,2,3,3)
Values = c(1,2,3,4,5,6,7,8,9)
tapply(Values,Id,sum)  
```

Sequence function:
A sequence is one of the components of a 'for loop'.
Syntax: seq(from, to, by, length)
from -> starting number
to -> ending number
by -> increment or decrement(width)
length -> number of elements required

```{r}
seq(from=1, to=10, by=2)
seq(from=10, to=100, length=4)
```

DATA VISUALIZATION IN R

Scatter Plot
```{r}
X = 1:10
Y = X^2
plot(Y)
```

Dataset 'mtcars': A data frame with 32 observations on 11 variables.

```{r}
plot(mtcars$wt, mtcars$mpg, main="ScatterPlot", xlab="Car weight", ylab="Miles per Gallon", pch=19)
```

Line Plot:
```{r}
X = 1:10
Y = X^2
plot(X,Y,type='l')
```

Bar Plot:
Syntax: barplot(list, names.arg, xlab, ylab, main, names.arg, col)
```{r}
H = c(7,12,28,3,41)
M = c("Mar", "Apr", "May", "Jun", "Jul")
barplot(H, names.arg = M, xlab = "Month", ylab = "Revenue", col = "blue", main = "Revenue chart", border = "blue")
```


