''' By Shubham Agarwal
    Link: https://github.com/BeAgarwal/R '''
  
BASIC OF R

To clear the variable environment.

```{r}
rm(list=ls())
```

To clear particular variable from environment.

```{r}
a=10
rm(a)
```

To clear the console.
control + L

We can define a variable using = and <-

```{r}
a = 10
b <- 20
print(a)
print(b)
```

To concat the variable.

```{r}
print(c(a,b))
```

To save a single variable available in workspace
& to save a full workspace with specified file name.

```{r}
save(a,file="file1.Rdata")
save(list=ls(all.names=TRUE),file="file1.Rdata")
save.image() #shortcut to save whole workspace
```

To load the saved workspace

```{r}
load(file="file1.Rdata")
```

Predefined constants in R:

```{r}
pi
letters
LETTERS
month.name
month.abb
```

Find data type of object

```{r}
typeof(a)
```

Verify if object is of certain datatype

```{r}
is.integer(a)
is.double(a)
```

Convert into another data type (Type Casting)

```{r}
as.integer(a)
as.complex(a)
```

VECTOR: Ordered collection of same data types
LIST: Ordered collection of objects
DATAFRAME: Generic tabular objects

Vector

```{r}
vec = c(1.2, 3.4 , 4.2)
print(vec)
```

List

```{r}
l = list(1, 2.7, "andrew")
print(l)
```

```{r}
l = list("Roll"=1,"Name"="Andrew")
print(l$Name)
```

```{r}
l = list(1,3,5,list(2,4,6),list(7,8,9))
l[[5]][[1]] #7
l[[3]]      #5
l[[4]][[3]] #6
```

Update in the list

```{r}
l[[2]] = "C2"
l[[2]]
```

Data Frame

```{r}
roll = c(1,2,3,4,5)
name = c("Andrew","Amit","Bob","Charles","George")
course = c("MCA1","MCA2","MCA1","MCA5","MCA3")
df = data.frame(roll,name,course)
df
```

Create a dataframe using data from a file

DF = read.table(path="",sep="")


subset() which extracts subset of data based on conditions.

```{r}
pd = data.frame("Name"=c("Andrew","Amit","Bob","Charles","George"),"roll" = c(1,2,3,4,5),"course" = c("MCA1","MCA2","MCA1","MCA5","MCA3"))

pd2 = subset(pd, roll>3)
print(pd2)
```

Adding extra rows and extra columns using rbind and cbind.

```{r}
df = rbind(pd, data.frame(Name = "Michael",roll = 6, course = "MCA6"))
print(df)

df = cbind(df, phone=c(123,456,789,321,654,987))
print(df)
```

Deleting rows and columns

```{r}
df
df = df[-6,]  #Deleting row
df
df = df[,!names(df) %in% c("phone")]   #Deleting column
df
df = df[!df$Name == "Charles",] #Condition based deletion
df
```

RECASTING
is the process of manipulating a data frame in terms of its variable.
We have to use the library reshape2.

```{r}
install.packages('reshape2')
```

Recast can be done in two steps.
-> Melt
melt(data, id.vars, measures.vars, variable.name="variable", value.name="value")

-> Cast
dcast(data, formula, value.var = col. with values)

```{r}
pd = data.frame("Name"=c("Andrew","Amit","Bob","Charles","George"),"Age"=c(25,21,18,26,28),"Month"=c("Jan","Feb","Jun","Dec","May"),"BP"=c(90,78,65,120,110))
print(pd)

library(reshape2)
meltpd = melt(pd, id.vars=c("Name","Month"), 
              measure.vars=c("Age","BP"))
print(meltpd)

dcastpd = dcast(meltpd, variable+Month ~ Name, value.var = "value")
print(dcastpd)
```

Recast in single step.
recast(data, formula, ....,id.var, measure.var)

```{r}
pd = data.frame("Name"=c("Andrew","Amit","Bob","Charles","George"),"Age"=c(25,21,18,26,28),"Month"=c("Jan","Feb","Jun","Dec","May"),"BP"=c(90,78,65,120,110))
print(pd)

recastpd = recast(pd, variable+Month ~ Name, id.var = c("Name","Month"))
print(recastpd)
```

library dplyr is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges: 
mutate() adds new variables that are functions of existing variables. 
select() picks variables based on their names.

```{r}
install.packages('dplyr')
```

```{r}
library(dplyr)
pd2 = mutate(pd, log_BP = log(BP))
print(pd2)

pd3 = select(pd, "Name", "BP")
print(pd3)
```

Joining of two data frames.
Follwing commands can be used to combine datasets.
-> left_join()
-> right_join()
-> full_join()
-> inner_join()
-> anti_join()

```{r}
library(dplyr)

df1 = data.frame("Roll"=c(1,2,3,4,5,6,7,8,9,10),"Name"=c("Andrew","Annie","Amit","Bob","Charles","Daizy","George","Mini","Luna","Kamal"))

df2 = data.frame("Roll"=c(1,2,3,4,5,6,7,8,9,10),"Gender"=c('M','F','M','M','M','F','M','F','F','M'))

df3 = data.frame("Roll"=c(1,2,3,4,5,6,7,8,9,10),"Marks"=c(98,65,78,34,67,81,49,56,45,90))

df_full_join = full_join(df2, df1, by="Roll")
df_full_join2 = full_join(df_full_join, df3, by="Roll")
print(df_full_join2)
```


Arithmetic:
Hierarchy of operations:
Bracket                    ()
Exponent                   ^, **
Division                    /
Multiplication              *
Addition and subtraction   +,-

Logical:
Symbols: <, <=, >, >=, ==, !=, !, |, &, isTRUE

Matrix Operations:
To create a matrix:


```{r}
A = matrix()
B = matrix(c(1,2,3,4,5,6,7,8,9))
C = matrix(c(1,2,3,4,5,6,7,8,9), nrow = 3, ncol = 3, byrow = FALSE)
D = matrix(3,3,4)
E = diag(c(4,5,6),3,3)  #DIAGONAL MATRIX
I = diag(1,3,3)     #IDENTITY MATRIX
```

MATRIX METRICS:

Finding the size of the matrix, A:
dim(A) will return the size of the matrix
nrow(A) will return the number of rows
ncol(A) will return the number of columns
prod(dim(A)) or length(A) will return the number of elements
```{r}
A = matrix(c(1,2,3,4,5,6,7,8,9), nrow = 3, ncol = 3, byrow = TRUE)
dim(A)
nrow(A)
ncol(A)
prod(dim(A))
```

Strings can be assigned as names of rows and columns using rownames() and colnames()

```{r}
rownames(A) = c("a","b","c")
colnames(A) = c("d","e","f")
A
```

Accessing from matrix:
```{r}
A
A[,1] #access the column
A[2,3]
A[2,]  #access the row
A[,-2] #access the column that has to be avoided
A[-2,] #access the row that has to be avoided
```

Colon Operator:
```{r}
1:10
10:-1
```

Colon notation can also be used to pick sub-matrices.
```{r}
A
A[1:3,1:2]
A[1:3,-3]
A[c(1,3),1:2]
```

Matric Concatenation: refers to merging of a row or column to a matrix
->Concatenation of a row to a matrix is done sing rbind()
->Concatenation of a column to a matrix is done using cbind()
->Consistency of the dimensions between the matrix and the vector should be checked before concatenation
```{r}
A
B = matrix(c(10,11,12), nrow = 1, ncol = 3)
C = rbind(A,B)
C
```

Deleting the column or row
```{r}
A = A[-4,]
A
```

Matrix Algebra:
Addition             A+B
Subtraction          A-B
Multiplication       
      Regular matrix multiplication  A%*%B
      Element-wise multiplication    A*B
Matrix Division      A/B (element-wise)

```{r}
mat1 = matrix(c(2,2,2,2), ncol = 2, nrow = 2)
mat1
mat2 = matrix(c(1,2,3,4), nrow = 2, ncol = 2)
mat2
mat1 + mat2
mat1 - mat2
mat1 %*% mat2
mat1 * mat2
mat1 / mat2
```









